---
title: "all plate analysis 15min-2h"
author: "Kara O'Driscoll"
date: "2024-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

WITHOUT N3

## opening correct files

```{r}
rm(list=objects())
library(tidyverse)
library(stringr)  
library(dplyr) 
library(ggfortify)
library(cowplot)

Path_to_working_directory <- "/Users/karaodriscoll/Documents/Lab Masters/all plates, 2h treated"
setwd(Path_to_working_directory) #sets directory to location of dataset
#opening dataset 
Name_of_dataset <- gsub(".*/","", Path_to_working_directory)
list_of_files <- list.files(pattern="*.csv")
raw_data <- read_csv(list_of_files, id = "file_name", show_col_types = FALSE)


#want to obtain the different time points for the treatment
raw_data_with_time_point <- raw_data %>%
  mutate(time_point = str_extract(file_name, "(?<=\\()\\d+(?=\\)\\(000\\)\\.csv)")) %>% #basically means that it looks at all the files with N2,000 and csv then labels by what is different in the name
 mutate(plate_number = str_extract(file_name, "(?<=N)\\d+"))

raw_data_with_time_point$time_point <- as.numeric(raw_data_with_time_point$time_point)
raw_data_with_time_point$plate_number <- as.numeric(raw_data_with_time_point$plate_number)
raw_data_with_time_point$time_point <- as.factor(raw_data_with_time_point$time_point)

```

1\.

# 1. number of spikes

```{r}
#NUMBER OF SPIKES
number_of_spikes_data <- raw_data_with_time_point %>%
  filter(Well.Averages=="Number of Spikes") %>%
  mutate(across(-c(1,2,27,28), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Number_of_Spikes")

#treatment data frame 
treatment <- raw_data_with_time_point %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#combining treatment and parameter - spikes
spikes_with_treatment_data <- left_join(number_of_spikes_data, treatment, by = c("file_name", "time_point", "well_ID"))

```

```{r}
#normalize - created dfs for the different timepoints
baseline_for_spikes <- spikes_with_treatment_data %>%
  filter(time_point==000) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(Baseline_spikes=Number_of_Spikes)

timepoint1_for_spikes <- spikes_with_treatment_data %>%
  filter(time_point==001) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Timepoint1_spikes=Number_of_Spikes)

timepoint2_for_spikes <- spikes_with_treatment_data %>% 
  filter(time_point==002)%>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Timepoint2_spikes=Number_of_Spikes)

timepoint3_for_spikes <- spikes_with_treatment_data %>%
  filter(time_point==003)%>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Timepoint3_spikes=Number_of_Spikes)

timepoint4_for_spikes <- spikes_with_treatment_data %>%
  filter(time_point==004)%>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Timepoint4_spikes=Number_of_Spikes)

timepoint5_for_spikes <- spikes_with_treatment_data %>%
  filter(time_point==005)%>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Timepoint5_spikes=Number_of_Spikes)

timepoint6_for_spikes <- spikes_with_treatment_data %>%
  filter(time_point==006) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Timepoint6_spikes=Number_of_Spikes)


#merging the new dfs 
merged_data <- merge(baseline_for_spikes, timepoint1_for_spikes, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data <- merge(merged_data, timepoint2_for_spikes, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data <- merge(merged_data, timepoint3_for_spikes, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data <- merge(merged_data, timepoint4_for_spikes, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data <- merge(merged_data, timepoint5_for_spikes, by = c("well_ID", "Treatment", "plate_number.x"))  
merged_data <- merge(merged_data, timepoint6_for_spikes, by = c("well_ID", "Treatment", "plate_number.x"))

#now normalising the data, had to create new DF so that each column was included

merged_data_normalised <- mutate(merged_data, "0"= merged_data$Baseline_spikes - merged_data$Baseline_spikes) 
merged_data_normalised. <- mutate(merged_data_normalised, "15"= Timepoint1_spikes - Baseline_spikes) 
merged_data_normalised.. <- mutate(merged_data_normalised., "35"= Timepoint2_spikes - Baseline_spikes)
merged_data_normalised... <- mutate(merged_data_normalised.., "55"= Timepoint3_spikes - Baseline_spikes)
merged_data_normalised__ <- mutate(merged_data_normalised..., "75"= Timepoint4_spikes - Baseline_spikes)
merged_data_normalised_. <- mutate(merged_data_normalised__, "95"= Timepoint5_spikes - Baseline_spikes)
merged_data_normalised._ <- mutate(merged_data_normalised_., "115"= Timepoint6_spikes - Baseline_spikes) 
spikes_normalised <- merged_data_normalised._ %>%
  select(-c(4:10)) %>%
  pivot_longer(cols = c(4:10), 
               names_to = "Timepoint", 
               values_to = "Normalised_data") #pivoting data allowed all the normalised data to be in one column for plotting 
spikes_normalised$Timepoint <- as.numeric(spikes_normalised$Timepoint) 
```

```{r}
spikes_normalised_avg <- spikes_normalised %>%
  group_by(Treatment, Timepoint) %>% 
  mutate(average = mean(Normalised_data), IQR = IQR(Normalised_data)) #IQR used for geom_ribbon function but doesn't actually look that good 

shapiro.test(spikes_normalised_avg$Normalised_data)

spikes_normalised_avg$Treatment <- as.factor(spikes_normalised_avg$Treatment)
spikes_normalised_avg$`plate_number.x` <- as.factor(spikes_normalised_avg$`plate_number.x`)
spikes_normalised_avg$Timepoint <- as.factor(spikes_normalised_avg$Timepoint)

#all have to be factor for ANOVA for DoF !! how excite 

#but then numeric for plotting Y=UGH

spikes_normalised_avg$Timepoint <- as.numeric(spikes_normalised_avg$Timepoint)
```

```{r}
#expeerimenting with stats - three way anova 
data_for_stats <- spikes_normalised_avg %>% 
  drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
  filter(is.finite(LogParameter))



model_treatment <- lm(`LogParameter`~Treatment*Timepoint, data = data_for_stats) 
#diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

```{r}
#post-hoc
TukeyHSD(aov(model_treatment), "Treatment", ordered = TRUE)


summary(aov(model_treatment), test = adjusted(type = "BH"))

```

2.  

    ```{r}
    spikes_normalised$Timepoint <- as.numeric(spikes_normalised$Timepoint) 

    spikes_normalised_avg <- spikes_normalised_avg %>%
      mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

    ggplot(spikes_normalised_avg, aes(x = Timepoint, colour = Treatment)) +
      geom_line(aes(y=average)) +
      geom_point(aes(y=Normalised_data, shape = Treatment)) +
      labs(title = "Normalised Number of Spikes by Treatment",
           x = "Time (mins)",
           y = "Normalised from baseline of Number of Spikes")+
      theme_cowplot()

    ggsave("spikes_normalised_avg.png")
    ```

    so this graph: normalised treatment to before treated timepoint then subtracted number of spikes from basline - dots

    anyway shows effect of treatment - 5meo compared to psilocin and vehicle but nothing else, not time

    will now plot not normalised :) - did this, in separate markdown just for NoS and found that raw data was not any sig, bit also all had different starting points but also looked like doi and 5meo dec over time wherreas psilocin inc?

    # 2. mean firing rate

```{r}
#MEAN FIRING RATE (Hz)
mean_firing_rate_data <- raw_data_with_time_point %>%
  filter(Well.Averages=="Mean Firing Rate (Hz)") %>%
  mutate(across(-c(1,2,27,28), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Mean Firing Rate (Hz)")

#if cleared, need to rerun treatment df too
#treatment data frame 
treatment <- raw_data_with_time_point %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")
#to join

mean_firing_rate_with_treatment_data <- left_join(mean_firing_rate_data, treatment, by = c("file_name", "time_point", "well_ID"))

```

```{r}
#normalize - created dfs for the different timepoints
baseline_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(time_point==000) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(Baseline_mfr=`Mean Firing Rate (Hz)`)

timepoint1_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(time_point==001) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint1_mfr=`Mean Firing Rate (Hz)`)

timepoint2_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(time_point==002) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint2_mfr=`Mean Firing Rate (Hz)`)

timepoint3_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(time_point==003) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint3_mfr=`Mean Firing Rate (Hz)`)

timepoint4_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(time_point==004) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint4_mfr=`Mean Firing Rate (Hz)`)

timepoint5_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(time_point==005) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint5_mfr=`Mean Firing Rate (Hz)`)

timepoint6_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(time_point==006) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint6_mfr=`Mean Firing Rate (Hz)`)

#merging the new dfs 
merged_data_mfr <- merge(baseline_for_mfr, timepoint1_for_mfr, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_mfr <- merge(merged_data_mfr, timepoint2_for_mfr, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_mfr <- merge(merged_data_mfr, timepoint3_for_mfr, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_mfr <- merge(merged_data_mfr, timepoint4_for_mfr, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_mfr <- merge(merged_data_mfr, timepoint5_for_mfr, by = c("well_ID", "Treatment", "plate_number.x"))  
merged_data_mfr <- merge(merged_data_mfr, timepoint6_for_mfr, by = c("well_ID", "Treatment", "plate_number.x"))

#now normalising the data

merged_data_normalised <- mutate(merged_data_mfr, "0"= Baseline_mfr - Baseline_mfr) 
merged_data_normalised. <- mutate(merged_data_normalised, "15"= timepoint1_mfr - Baseline_mfr) 
merged_data_normalised.. <- mutate(merged_data_normalised., "35"= timepoint2_mfr - Baseline_mfr)
merged_data_normalised... <- mutate(merged_data_normalised.., "55"= timepoint3_mfr - Baseline_mfr)
merged_data_normalised__ <- mutate(merged_data_normalised..., "75"= timepoint4_mfr - Baseline_mfr)
merged_data_normalised_. <- mutate(merged_data_normalised__, "95"= timepoint5_mfr - Baseline_mfr)
merged_data_normalised._ <- mutate(merged_data_normalised_., "115"= timepoint6_mfr - Baseline_mfr) 
mfr_normalised <- merged_data_normalised._ %>%
  select(-c(4:10)) %>%
  pivot_longer(cols = c(4:10), 
               names_to = "Timepoint", 
               values_to = "Normalised_data")
mfr_normalised$Timepoint <- as.numeric(mfr_normalised$Timepoint)

```

```{r}
#averaging data
mfr_normalised_avg <- mfr_normalised %>%
  group_by(Treatment, Timepoint) %>% 
  mutate(average = mean(Normalised_data), IQR = IQR(Normalised_data))

shapiro.test(mfr_normalised_avg$Normalised_data)

mfr_normalised_avg$Treatment <- as.factor(mfr_normalised_avg$Treatment)
mfr_normalised_avg$`plate_number.x` <- as.factor(mfr_normalised_avg$`plate_number.x`)

mfr_normalised_avg$Timepoint <- as.factor(mfr_normalised_avg$Timepoint)
```

```{r}
#log 
data_for_stats <- mfr_normalised_avg %>% 
  drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Timepoint, data = data_for_stats) 
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

```{r}
#post-hoc

TukeyHSD(aov(model_treatment), "Treatment", ordered = TRUE)


summary(aov(model_treatment), test = adjusted(type = "BH"))
```

also shows same as NoS

```{r}
mfr_normalised_avg$Timepoint <- as.numeric(mfr_normalised_avg$Timepoint)

mfr_normalised_avg <- mfr_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

ggplot(mfr_normalised_avg, aes(x = Timepoint, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Mean Firing Rate by Treatment",
       x = "Time (mins)",
       y = "Normalised from baseline of Mean Firing Rate (Hz)")+
  theme_cowplot()

ggsave("mfr_normalised_avg.png")
```

ok once again: line is average, averaged by time and treatment then the dots show the individual wells ! - this is the normalised not log plotted because the log removes values of 0...

# 3. burst frequency

```{r}
#extract parameter
burst_frequency_data <- raw_data_with_time_point %>%
  filter(Well.Averages=="Burst Frequency - Avg (Hz)") %>%
  mutate(across(-c(1,2,27,28), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Burst_Frequency_(Hz)")

#extract treatment
treatment <- raw_data_with_time_point %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#joining
burst_with_treatment <- left_join(burst_frequency_data, treatment, by = c("file_name", "time_point", "well_ID"))

```

```{r}
baseline_for_burst <- burst_with_treatment %>%
  filter(time_point==000) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(Baseline_burst=`Burst_Frequency_(Hz)`)

timepoint1_for_burst <- burst_with_treatment %>%
  filter(time_point==001) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint1_burst=`Burst_Frequency_(Hz)`)

timepoint2_for_burst <- burst_with_treatment %>%
  filter(time_point==002) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint2_burst=`Burst_Frequency_(Hz)`)

timepoint3_for_burst <- burst_with_treatment %>%
  filter(time_point==003) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint3_burst=`Burst_Frequency_(Hz)`)

timepoint4_for_burst <- burst_with_treatment %>%
  filter(time_point==004) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint4_burst=`Burst_Frequency_(Hz)`)

timepoint5_for_burst <- burst_with_treatment %>%
  filter(time_point==005) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint5_burst=`Burst_Frequency_(Hz)`)

timepoint6_for_burst <- burst_with_treatment %>%
  filter(time_point==006) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point",  "plate_number.y")) %>%
  rename(timepoint6_burst=`Burst_Frequency_(Hz)`)

#then need to combine them / MERGE
merged_data_burst <- merge(baseline_for_burst, timepoint1_for_burst, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_burst <- merge(merged_data_burst, timepoint2_for_burst, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_burst <- merge(merged_data_burst, timepoint3_for_burst, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_burst <- merge(merged_data_burst, timepoint4_for_burst, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_burst <- merge(merged_data_burst, timepoint5_for_burst, by = c("well_ID", "Treatment", "plate_number.x"))  
merged_data_burst <- merge(merged_data_burst, timepoint6_for_burst, by = c("well_ID", "Treatment", "plate_number.x"))

#normalise
merged_data_normalised <- mutate(merged_data_burst, "0"= Baseline_burst - Baseline_burst) 
merged_data_normalised. <- mutate(merged_data_normalised, "15"= timepoint1_burst - Baseline_burst) 
merged_data_normalised.. <- mutate(merged_data_normalised., "35"= timepoint2_burst - Baseline_burst)
merged_data_normalised... <- mutate(merged_data_normalised.., "55"= timepoint3_burst - Baseline_burst)
merged_data_normalised__ <- mutate(merged_data_normalised..., "75"= timepoint4_burst - Baseline_burst)
merged_data_normalised_. <- mutate(merged_data_normalised__, "95"= timepoint5_burst - Baseline_burst)
merged_data_normalised._ <- mutate(merged_data_normalised_., "115"= timepoint6_burst - Baseline_burst) 

#pivot longer - so only normalised data is there
burst_normalised <- merged_data_normalised._ %>%
  select(-c(4:10)) %>%
  pivot_longer(cols = c(4:10), 
               names_to = "Timepoint", 
               values_to = "Normalised_data")

#make numeric so can geom_line
burst_normalised$Timepoint <- as.numeric(burst_normalised$Timepoint)

#remove NA
burst_normalised <- na.omit(burst_normalised)

```

```{r}
#average
bf_avg <- burst_normalised %>%
  group_by(Treatment, Timepoint) %>% 
  mutate(average = mean(Normalised_data))

shapiro.test(bf_avg$Normalised_data)
bf_avg$Treatment <- as.factor(bf_avg$Treatment)
bf_avg$`plate_number.x` <- as.factor(bf_avg$`plate_number.x`)
bf_avg$Timepoint <- as.factor(bf_avg$Timepoint)

bf_avg$Timepoint <- as.numeric(bf_avg$Timepoint)
```

```{r}
#statssss
data_for_stats <- bf_avg %>% 
  drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Timepoint, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

in this case - no significance !

```{r}

bf_avg$Timepoint <- as.numeric(bf_avg$Timepoint)

bf_avg <- bf_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

ggplot(bf_avg, aes(x = Timepoint, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Burst Frequency by Treatment",
       x = "Time (mins)",
       y = "Normalised from baseline of Burst Frequency (Hz)")+
  theme_cowplot()

ggsave("bf_avg.png")
```

so no significance! not that many bursts / that strong either - so would say treatment doesnt increase amount / strength of bursts at 2h .

# 4. spikes per burst

```{r}
number_of_spikes_per_burst_data <- raw_data_with_time_point %>%
  filter(Well.Averages=="Number of Spikes per Burst - Avg") %>%
  mutate(across(-c(1,2,27,28), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "spikes_per_burst")

#extract treatment
treatment <- raw_data_with_time_point %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#joining 
spikes_per_burst_with_treatment_data <- left_join(number_of_spikes_per_burst_data, treatment, by = c("file_name", "time_point", "well_ID"))

```

```{r}
#normalise

baseline_for_spb <- spikes_per_burst_with_treatment_data %>%
  filter(time_point==000) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Baseline_spb=`spikes_per_burst`)

timepoint1_for_spb <- spikes_per_burst_with_treatment_data %>%
  filter(time_point==001) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint1_spb=`spikes_per_burst`)

timepoint2_for_spb <- spikes_per_burst_with_treatment_data %>%
  filter(time_point==002) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint2_spb=`spikes_per_burst`)

timepoint3_for_spb <- spikes_per_burst_with_treatment_data %>%
  filter(time_point==003) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint3_spb=`spikes_per_burst`)

timepoint4_for_spb <- spikes_per_burst_with_treatment_data %>%
  filter(time_point==004) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint4_spb=`spikes_per_burst`)

timepoint5_for_spb <- spikes_per_burst_with_treatment_data %>%
  filter(time_point==005) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint5_spb=`spikes_per_burst`)

timepoint6_for_spb <- spikes_per_burst_with_treatment_data %>%
  filter(time_point==006) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint6_spb=`spikes_per_burst`)

#then need to combine them / MERGE
merged_data_spb <- merge(baseline_for_spb, timepoint1_for_spb, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_spb <- merge(merged_data_spb, timepoint2_for_spb, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_spb <- merge(merged_data_spb, timepoint3_for_spb, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_spb <- merge(merged_data_spb, timepoint4_for_spb, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_spb <- merge(merged_data_spb, timepoint5_for_spb, by = c("well_ID", "Treatment", "plate_number.x"))  
merged_data_spb <- merge(merged_data_spb, timepoint6_for_spb, by = c("well_ID", "Treatment", "plate_number.x"))

#normalise
merged_data_normalised <- mutate(merged_data_spb, "0"= Baseline_spb - Baseline_spb) 
merged_data_normalised. <- mutate(merged_data_normalised, "15"= timepoint1_spb - Baseline_spb) 
merged_data_normalised.. <- mutate(merged_data_normalised., "35"= timepoint2_spb - Baseline_spb)
merged_data_normalised... <- mutate(merged_data_normalised.., "55"= timepoint3_spb - Baseline_spb)
merged_data_normalised__ <- mutate(merged_data_normalised..., "75"= timepoint4_spb - Baseline_spb)
merged_data_normalised_. <- mutate(merged_data_normalised__, "95"= timepoint5_spb - Baseline_spb)
merged_data_normalised._ <- mutate(merged_data_normalised_., "115"= timepoint6_spb - Baseline_spb) 

#pivot longer - so only normalised data is there
spb_normalised <- merged_data_normalised._ %>%
  select(-c(4:10)) %>%
  pivot_longer(cols = c(4:10), 
               names_to = "Timepoint", 
               values_to = "Normalised_data")

#make numeric so can geom_line
spb_normalised$Timepoint <- as.numeric(spb_normalised$Timepoint)

#remove NA
spb_normalised <- na.omit(spb_normalised)

```

```{r}
spb_normalised_avg <- spb_normalised %>%
  group_by(Treatment, Timepoint) %>% 
  mutate(average = mean(Normalised_data), IQR = IQR(Normalised_data))

shapiro.test(spb_normalised_avg$Normalised_data)

spb_normalised_avg$Treatment <- as.factor(spb_normalised_avg$Treatment)
spb_normalised_avg$`plate_number.x` <- as.factor(spb_normalised_avg$`plate_number.x`)
spb_normalised_avg$Timepoint <- as.factor(spb_normalised_avg$Timepoint)

spb_normalised_avg$Timepoint <- as.numeric(spb_normalised_avg$Timepoint)
```

```{r}
data_for_stats <- spb_normalised_avg %>% 
  drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Timepoint, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

also no sig!

so single-electrode bursting not affected by treatment! in first 2h

```{r}
spb_normalised_avg$Timepoint <- as.numeric(spb_normalised_avg$Timepoint)

spb_normalised_avg <- spb_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

ggplot(spb_normalised_avg, aes(x = Timepoint, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Spikes per Burst by Treatment",
       x = "Time (mins)",
       y = "Normalised from baseline of Spikes per Burst")+
  theme_cowplot()

ggsave("spb_normalised_avg.png")
```

# 5. Synchrony Index

```{r}
#SYNCHRONY INDEX
#extracting parameter
synchrony_index_data <- raw_data_with_time_point %>%
  filter(Well.Averages=="Synchrony Index") %>%
  mutate(across(-c(1,2,27,28), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Synchrony Index")

#extract treatment
treatment <- raw_data_with_time_point %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#joining 
synchrony_index_with_treatment_data <- left_join(synchrony_index_data, treatment, by = c("file_name", "time_point", "well_ID"))
```

\

```{r}
#Creates new df for the different timepoints 
baseline_for_SI <- synchrony_index_with_treatment_data %>%
  filter(time_point==000) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(Baseline_si=`Synchrony Index`)

timepoint1_for_SI <- synchrony_index_with_treatment_data %>%
  filter(time_point==001) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint1_si=`Synchrony Index`)

timepoint2_for_SI <- synchrony_index_with_treatment_data %>%
  filter(time_point==002) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint2_si=`Synchrony Index`)

timepoint3_for_SI <- synchrony_index_with_treatment_data %>%
  filter(time_point==003) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint3_si=`Synchrony Index`)

timepoint4_for_SI <- synchrony_index_with_treatment_data %>%
  filter(time_point==004) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint4_si=`Synchrony Index`)

timepoint5_for_SI <- synchrony_index_with_treatment_data %>%
  filter(time_point==005) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint5_si=`Synchrony Index`)

timepoint6_for_SI <- synchrony_index_with_treatment_data %>%
  filter(time_point==006) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "time_point", "plate_number.y")) %>%
  rename(timepoint6_si=`Synchrony Index`)

#then need to combine them / MERGE
merged_data_si <- merge(baseline_for_SI, timepoint1_for_SI, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_si <- merge(merged_data_si, timepoint2_for_SI, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_si <- merge(merged_data_si, timepoint3_for_SI, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_si <- merge(merged_data_si, timepoint4_for_SI, by = c("well_ID", "Treatment", "plate_number.x"))
merged_data_si <- merge(merged_data_si, timepoint5_for_SI, by = c("well_ID", "Treatment", "plate_number.x"))  
merged_data_si <- merge(merged_data_si, timepoint6_for_SI, by = c("well_ID", "Treatment", "plate_number.x"))

#normalise
merged_data_normalised <- mutate(merged_data_si, "0"= Baseline_si - Baseline_si) 
merged_data_normalised. <- mutate(merged_data_normalised, "15"= timepoint1_si - Baseline_si) 
merged_data_normalised.. <- mutate(merged_data_normalised., "35"= timepoint2_si - Baseline_si)
merged_data_normalised... <- mutate(merged_data_normalised.., "55"= timepoint3_si - Baseline_si)
merged_data_normalised__ <- mutate(merged_data_normalised..., "75"= timepoint4_si - Baseline_si)
merged_data_normalised_. <- mutate(merged_data_normalised__, "95"= timepoint5_si - Baseline_si)
merged_data_normalised._ <- mutate(merged_data_normalised_., "115"= timepoint6_si - Baseline_si) 

#pivot longer - so only normalised data is there
SI_normalised <- merged_data_normalised._ %>%
  select(-c(4:10)) %>%
  pivot_longer(cols = c(4:10), 
               names_to = "Timepoint", 
               values_to = "Normalised_data")

#make numeric so can geom_line
SI_normalised$Timepoint <- as.numeric(SI_normalised$Timepoint)

#remove NA
 SI_normalised <- na.omit(SI_normalised)
```

```{r}
SI_normalised_avg <- SI_normalised %>%
  group_by(Treatment, Timepoint) %>% 
  mutate(average = mean(Normalised_data))

shapiro.test(SI_normalised_avg$Normalised_data)

SI_normalised_avg$Treatment <- as.factor(SI_normalised_avg$Treatment)
SI_normalised_avg$`plate_number.x` <- as.factor(SI_normalised_avg$`plate_number.x`)
SI_normalised_avg$Timepoint <- as.factor(SI_normalised_avg$Timepoint)

SI_normalised_avg$Timepoint <- as.numeric(SI_normalised_avg$Timepoint)
```

```{r}
data_for_stats <- SI_normalised_avg %>% 
 drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Timepoint, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

```{r}
TukeyHSD(aov(model_treatment), "Treatment", ordered = TRUE)

summary(aov(model_treatment), test = adjusted(type = "BH"))

```

sig treatment between psilocin and vehicle

```{r}
SI_normalised_avg$Timepoint <- as.numeric(SI_normalised_avg$Timepoint)

SI_normalised_avg <- SI_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

ggplot(SI_normalised_avg, aes(x = Timepoint, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Synchrony Index by Treatment",
       x = "Time (mins)",
       y = "Normalised from baseline of Synchrony Index")+
  theme_cowplot()
```

maybe psilocin slighlty more inc??
