---
title: "fold-change for 2h"
author: "Kara O'Driscoll"
date: "2024-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## opening files

```{r}
rm(list=objects())
library(tidyverse)
library(stringr)  
library(dplyr) 
library(ggfortify)
library(cowplot)
library(readr)
library(RColorBrewer)


Path_to_working_directory <- "/Users/karaodriscoll/Documents/Lab Masters/fold change 2h"
setwd(Path_to_working_directory) #sets directory to location of dataset
#opening dataset 
Name_of_dataset <- gsub(".*/","", Path_to_working_directory)
list_of_files <- list.files(pattern="*.csv")
raw_data <- read_csv(list_of_files, id = "file_name", show_col_types = FALSE)

raw_data$Treatment <- as.factor(raw_data$Treatment) #need to make a factor to plot individual plate colours !!!!
raw_data$Plate <- as.factor(raw_data$Plate)

raw_data$Time <- as.factor(raw_data$Time)

raw_data <- raw_data %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

```

# inserting raincloud plot \<3

```{r}
packages <- c("cowplot", "readr", "ggplot2", "dplyr", "lavaan", "Hmisc")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))
}

# Load packages ----
library(ggplot2)

# Defining the geom_flat_violin function ----
# Note: the below code modifies the
# existing github page by removing a parenthesis in line 50

"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}

geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
          setup_data = function(data, params) {
            data$width <- data$width %||%
              params$width %||% (resolution(data$x, FALSE) * 0.9)
            
            # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
            data %>%
              group_by(group) %>%
              mutate(
                ymin = min(y),
                ymax = max(y),
                xmin = x,
                xmax = x + width / 2
              )
          },
          
          draw_group = function(data, panel_scales, coord) {
            # Find the points for the line to go all the way around
            data <- transform(data,
                              xminv = x,
                              xmaxv = x + violinwidth * (xmax - x)
            )
            
            # Make sure it's sorted properly to draw the outline
            newdata <- rbind(
              plyr::arrange(transform(data, x = xminv), y),
              plyr::arrange(transform(data, x = xmaxv), -y)
            )
            
            # Close the polygon: set first and last point the same
            # Needed for coord_polar and such
            newdata <- rbind(newdata, newdata[1, ])
            
            ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
          },
          
          draw_key = draw_key_polygon,
          
          default_aes = aes(
            weight = 1, colour = "grey20", fill = "white", size = 0.5,
            alpha = NA, linetype = "solid"
          ),
          
          required_aes = c("x", "y")
  )
```

# number of spikes

looking at raw data - stats

```{r}
shapiro.test(raw_data$`Number of Spikes`)
```

= not normally distributed so will convert to log

```         
```

```{r}

raw_data <- subset(raw_data, raw_data$`Number of Spikes` != 0) #removing NA and values of 0
data_for_stats <- raw_data %>% 
  drop_na(`Number of Spikes`) %>%
  mutate("LogParameter"=log(`Number of Spikes`))

model_treatment <- lm(`LogParameter` ~ Treatment * Time, data = data_for_stats) 
 #diagnostic plots

Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

no sig!

```{r}
ggplot(raw_data, aes(x = Time)) +
  
  geom_point(aes(y = `Number of Spikes`, colour = Treatment), position = position_jitter(width = .35))
```

not really sure what im trying to show here - that they are all at different points??

fold change

1.  JUST A boxplot - fold change was done by dividing well 2h after treatment by the same well before treatment, grouped by plate and condition!

plotting fold change data

```{r}
ggplot(raw_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_spikes`, colour = Treatment)) 
```

just plotting in a different way - raincloud plot

```{r}
ggplot(raw_data, aes(x = `Treatment`, y = `fold_change_spikes`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_spikes`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_spikes`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Number of Spikes')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  
  guides(fill = FALSE)+
  ggtitle('Number of Spikes 2h treated')
```

```{r}
model_treatment.1 <- lm(`fold_change_spikes`~ Treatment , data = raw_data) 
 #diagnostic plots

Diagnostic_plot.1 <- autoplot(model_treatment.1, smooth.colour = NA) 
print(Diagnostic_plot.1)


anova(model_treatment.1)

summary(aov(model_treatment.1))
```

no significance!

# mean firing rate

raw data

```{r}
shapiro.test(raw_data$`Mean Firing Rate (Hz)`) 
```

```{r}
raw_data <- subset(raw_data, raw_data$`Mean Firing Rate (Hz)` != 0) #removing NA and values of 0
data_for_stats.2 <- raw_data %>% 
  drop_na(`Mean Firing Rate (Hz)`) %>%
  mutate("LogParameter"=log(`Mean Firing Rate (Hz)`))

model_treatment.2 <- lm(`LogParameter` ~ Treatment, data = data_for_stats.2) 
 #diagnostic plots

Diagnostic_plot.2 <- autoplot(model_treatment.2, smooth.colour = NA) 
print(Diagnostic_plot.2)


anova(model_treatment.2)

summary(aov(model_treatment.2))
```

interaction

post hoc

```{r}
ggplot(raw_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_mfr`, colour = Treatment)) 
```

```{r}
ggplot(raw_data, aes(x = `Treatment`, y = `fold_change_mfr`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_mfr`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_mfr`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Mean Firing Rate')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  
  guides(fill = FALSE)+
  ggtitle('Mean Firing Rate 2h treated')
```

MFR and Number of Spikes are measuring the same thing just in slightly different ways - firing rate is Hz so based on frequency of the spikes, so graphs show basically same thing!

so with the fold change there is no sig diff of treatment on number of spikes over 2h. or mfr \>

```{r}
model_treatment.3 <- lm(`fold_change_mfr`~ Treatment, data = raw_data) 
 #diagnostic plots

Diagnostic_plot.3 <- autoplot(model_treatment.3, smooth.colour = NA) 
print(Diagnostic_plot.3)


anova(model_treatment.3)

summary(aov(model_treatment.3))
```

# spikes per burst

raw data

```{r}
shapiro.test(raw_data$`Number of Spikes per Burst - Avg`) 
```

```{r}
raw_data <- subset(raw_data, raw_data$`Number of Spikes per Burst - Avg` != 0) #removing NA and values of 0
data_for_stats.5 <- raw_data %>% 
  drop_na(`Number of Spikes per Burst - Avg`) %>%
  mutate("LogParameter"=log(`Number of Spikes per Burst - Avg`))

model_treatment.5 <- lm(`LogParameter` ~ Treatment * Time , data = data_for_stats.5) 
 #diagnostic plots

Diagnostic_plot.5 <- autoplot(model_treatment.5, smooth.colour = NA) 
print(Diagnostic_plot.5)

anova(model_treatment.5)

summary(aov(model_treatment.5))
```

newww sig

fold change

```{r}
spb_data <-  raw_data %>%
  drop_na(`fold_change_spb`)
```

```{r}
ggplot(spb_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_spb`, colour = Treatment)) 
```

doi really dec

```{r}
ggplot(spb_data, aes(x = `Treatment`, y = `fold_change_spb`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_spb`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_spb`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Spikes per Burst')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Spikes per Burst 2h treated')
```

not every well have a spike per burst - nor does every condition

```{r}
model_treatment.4 <- lm(`fold_change_spb`~ Treatment , data = spb_data) 
 #diagnostic plots

#wouldnt let me plot so just did ANOVA 

anova(model_treatment.4)

summary(aov(model_treatment.4))
```

# burst frequency

raw data

```{r}
shapiro.test(raw_data$`Burst Frequency - Avg (Hz)`)
```

```{r}
raw_data <- subset(raw_data, raw_data$`Burst Frequency - Avg (Hz)` != 0) #removing NA and values of 0
data_for_stats.7 <- raw_data %>% 
  drop_na(`Burst Frequency - Avg (Hz)`) %>%
  mutate("LogParameter"=log(`Burst Frequency - Avg (Hz)`))

model_treatment.7 <- lm(`LogParameter` ~ Treatment * Time, data = data_for_stats.7) 
 #diagnostic plots

Diagnostic_plot.7 <- autoplot(model_treatment.7, smooth.colour = NA) 
print(Diagnostic_plot.7)

anova(model_treatment.7)

summary(aov(model_treatment.7))

```

now nooo sig

fold change

```{r}
bf_data <- raw_data %>%
  drop_na(`fold_change_burst`)
```

```{r}
ggplot(bf_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_burst`, colour = Treatment)) 
```

```{r}
ggplot(bf_data, aes(x = `Treatment`, y = `fold_change_burst`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_burst`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_burst`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Burst Frequency')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Burst Frequency 2h treated')
```

also not every well has a burst - as in plates are different for different conditions, 1 and 3 have no bursts for 5MEO ??

```{r}
model_treatment.6 <- lm(`fold_change_burst`~ Treatment , data = bf_data) 
 #diagnostic plots


anova(model_treatment.6)

summary(aov(model_treatment.6))
```

# synchrony index

raw data

```{r}

shapiro.test(raw_data$`Synchrony Index`) 
```

```{r}
raw_data <- subset(raw_data, raw_data$`Synchrony Index` != 0) #removing NA and values of 0
data_for_stats.8 <- raw_data %>% 
  drop_na(`Synchrony Index`) %>%
  mutate("LogParameter"=log(`Synchrony Index`))

model_treatment.8 <- lm(`LogParameter` ~ Treatment  * Time, data = data_for_stats.8) 
 #diagnostic plots

Diagnostic_plot.8 <- autoplot(model_treatment.8, smooth.colour = NA) 
print(Diagnostic_plot.8)

anova(model_treatment.8)

summary(aov(model_treatment.8))
```

now no sig

fold change

```{r}
si_data <- raw_data %>%
  drop_na(`fold_change_si`)
```

```{r}
ggplot(si_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_si`, colour = `Treatment`)) 
```

not much change, not very sync at 2h

```{r}
ggplot(si_data, aes(x = `Treatment`, y = `fold_change_si`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_si`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_si`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Synchrony Index')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Synchrony Index 2h treated')
```

```{r}
model_treatment.9 <- lm(`fold_change_si`~ Treatment  , data = si_data) 
 #diagnostic plots

Diagnostic_plot.9 <- autoplot(model_treatment.9, smooth.colour = NA) 
print(Diagnostic_plot.9)

anova(model_treatment.9)

summary(aov(model_treatment.9))
```

NO SIG - NO INC IN SYNC AT 2H!
