---
title: "div25 fold change"
author: "Kara O'Driscoll"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## opening files

I removed from the excel file data with no active electrodes / values of 0

```{r}
rm(list=objects())
library(tidyverse)
library(stringr)  
library(dplyr) 
library(ggfortify)
library(cowplot)
library(readr)
library(RColorBrewer)


Path_to_working_directory <- "/Users/karaodriscoll/Documents/Lab Masters/fold change div25"
setwd(Path_to_working_directory) #sets directory to location of dataset
#opening dataset 
Name_of_dataset <- gsub(".*/","", Path_to_working_directory)
list_of_files <- list.files(pattern="*.csv")
raw_data <- read_csv(list_of_files, id = "file_name", show_col_types = FALSE)

raw_data$Treatment <- as.factor(raw_data$Treatment) #need to make a factor to plot individual plate colours !!!!
raw_data$Plate <- as.factor(raw_data$Plate)

raw_data$Time <- as.factor(raw_data$Time)

raw_data <- raw_data %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))
```

# insert raincloud plot

```{r}
packages <- c("cowplot", "readr", "ggplot2", "dplyr", "lavaan", "Hmisc")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))
}

# Load packages ----
library(ggplot2)

# Defining the geom_flat_violin function ----
# Note: the below code modifies the
# existing github page by removing a parenthesis in line 50

"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}

geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
          setup_data = function(data, params) {
            data$width <- data$width %||%
              params$width %||% (resolution(data$x, FALSE) * 0.9)
            
            # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
            data %>%
              group_by(group) %>%
              mutate(
                ymin = min(y),
                ymax = max(y),
                xmin = x,
                xmax = x + width / 2
              )
          },
          
          draw_group = function(data, panel_scales, coord) {
            # Find the points for the line to go all the way around
            data <- transform(data,
                              xminv = x,
                              xmaxv = x + violinwidth * (xmax - x)
            )
            
            # Make sure it's sorted properly to draw the outline
            newdata <- rbind(
              plyr::arrange(transform(data, x = xminv), y),
              plyr::arrange(transform(data, x = xmaxv), -y)
            )
            
            # Close the polygon: set first and last point the same
            # Needed for coord_polar and such
            newdata <- rbind(newdata, newdata[1, ])
            
            ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
          },
          
          draw_key = draw_key_polygon,
          
          default_aes = aes(
            weight = 1, colour = "grey20", fill = "white", size = 0.5,
            alpha = NA, linetype = "solid"
          ),
          
          required_aes = c("x", "y")
  )
```

# number of spikes

```{r}
shapiro.test(raw_data$`Number of Spikes`)
```

not normally distributed

```{r}
raw_data.n <- subset(raw_data, raw_data$`Number of Spikes` != 0) #removing NA and values of 0
data_for_stats <- raw_data.n %>% 
  drop_na(`Number of Spikes`) %>%
  mutate("LogParameter"=log(`Number of Spikes`))

model_treatment <- lm(`LogParameter` ~ Treatment * Time, data = data_for_stats) 
 #diagnostic plots

Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

time sig but not treatment

```{r}
ggplot(raw_data, aes(x = Time)) +
  geom_boxplot(aes(y= `Number of Spikes`, colour = Treatment))+
  geom_point(aes(y = `Number of Spikes`, colour = Treatment), position = position_jitter(width = .35))
```

kind of looks like an inc but not by much ?

fold change

```{r}
ggplot(raw_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_spikes`, colour = Treatment))
```

```{r}

ggplot(raw_data.n, aes(x = `Treatment`, y = `fold_change_spikes`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_spikes`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_spikes`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Number of Spikes')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  
  guides(fill = FALSE)+
  ggtitle('Number of Spikes 72h treated')

ggsave("raw_data.n.png")
```

stats

```{r}
model_treatment.1 <- lm(`fold_change_spikes`~ Treatment  , data = raw_data) 
 #diagnostic plots

Diagnostic_plot.1 <- autoplot(model_treatment.1, smooth.colour = NA) 
print(Diagnostic_plot.1)


anova(model_treatment.1)

summary(aov(model_treatment.1))
```

no sig diff!

# mean firing rate

```{r}

shapiro.test(raw_data$`Mean Firing Rate (Hz)`) 
```

not normally dis = to log

```{r}
raw_data.m <- subset(raw_data, raw_data$`Mean Firing Rate (Hz)` != 0) #removing NA and values of 0
data_for_stats.2 <- raw_data.m %>% 
  drop_na(`Mean Firing Rate (Hz)`) %>%
  mutate("LogParameter"=log(`Mean Firing Rate (Hz)`))

model_treatment.2 <- lm(`LogParameter` ~ Treatment  , data = data_for_stats.2) 
 #diagnostic plots

Diagnostic_plot.2 <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot.2)


anova(model_treatment.2)

summary(aov(model_treatment.2))
```

no sig

fold change

```{r}
ggplot(raw_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_mfr`, colour = Treatment)) 
```

compared to before treated, only one whcihi has really changed is doi - for plate 2 but not others, then psilocin for plate 3, but once again no difference/ not a general trend of the other plates

```{r}
ggplot(raw_data.m, aes(x = `Treatment`, y = `fold_change_mfr`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_mfr`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_mfr`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Mean Firing Rate')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  
  guides(fill = FALSE)+
  ggtitle('Mean Firing Rate 72h treated')

ggsave("raw_data.m.png")
```

stats

```{r}
model_treatment.3 <- lm(`fold_change_mfr`~ Treatment , data = raw_data) 
 #diagnostic plots

Diagnostic_plot.3 <- autoplot(model_treatment.3, smooth.colour = NA) 
print(Diagnostic_plot.3)


anova(model_treatment.3)

summary(aov(model_treatment.3))
```

no sig of anything

# spikes per burst

raw data looking at SPB

```{r}
shapiro.test(raw_data$`Number of Spikes per Burst - Avg`) 
```

not normally dis

```{r}
raw_data.s <- subset(raw_data, raw_data$`Number of Spikes per Burst - Avg` != 0) #removing NA and values of 0
data_for_stats.5 <- raw_data.s %>% 
  drop_na(`Number of Spikes per Burst - Avg`) %>%
  mutate("LogParameter"=log(`Number of Spikes per Burst - Avg`))

model_treatment.5 <- lm(`LogParameter` ~ Treatment * Time , data = data_for_stats.5) 
 #diagnostic plots

Diagnostic_plot.5 <- autoplot(model_treatment.5, smooth.colour = NA) 
print(Diagnostic_plot.5)


anova(model_treatment.5)

summary(aov(model_treatment.5))
```

no sig

looking at fold change

```{r}
spb_data <-  raw_data %>%
  drop_na(`fold_change_spb`)
```

```{r}
ggplot(spb_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_spb`, colour = Treatment)) 
```

```{r}
ggplot(spb_data, aes(x = `Treatment`, y = `fold_change_spb`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_spb`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_spb`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Spikes per Burst')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Spikes per Burst 72h treated')

ggsave("spb_data.png")
```

```{r}
model_treatment.4 <- lm(`fold_change_spb`~ Treatment , data = spb_data) 
 #diagnostic plots

Diagnostic_plot.4 <- autoplot(model_treatment.4, smooth.colour = NA) 
print(Diagnostic_plot.4)


anova(model_treatment.4)

summary(aov(model_treatment.4))
```

noooo sig anyway

# burst frequency

raw data

```{r}
shapiro.test(raw_data$`Burst Frequency - Avg (Hz)`) 
```

not normally dis = log

```{r}
raw_data.b <- subset(raw_data, raw_data$`Burst Frequency - Avg (Hz)` != 0) #removing NA and values of 0
data_for_stats.7 <- raw_data.b %>% 
  drop_na(`Burst Frequency - Avg (Hz)`) %>%
  mutate("LogParameter"=log(`Burst Frequency - Avg (Hz)`))

model_treatment.7 <- lm(`LogParameter` ~ Treatment * Time, data = data_for_stats.7) 
 #diagnostic plots

Diagnostic_plot.7 <- autoplot(model_treatment.7, smooth.colour = NA) 
print(Diagnostic_plot.7)


anova(model_treatment.7)

summary(aov(model_treatment.7))
```

nooo sig diff

fold change

```{r}
bf_data <- raw_data %>%
  drop_na(`fold_change_bursts`)
```

```{r}
ggplot(bf_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_bursts`, colour = Treatment)) 
```

```{r}
ggplot(bf_data, aes(x = `Treatment`, y = `fold_change_bursts`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_bursts`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_bursts`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Burst Frequency')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Burst Frequency 72h treated')

ggsave("bf_data.png")
```

```{r}
model_treatment.6 <- lm(`fold_change_bursts`~ Treatment  , data = bf_data) 
 #diagnostic plots

Diagnostic_plot.6 <- autoplot(model_treatment.6, smooth.colour = NA) 
print(Diagnostic_plot.6)


anova(model_treatment.6)

summary(aov(model_treatment.6))
```

also no sig

# synchrony index

raw data

```{r}
shapiro.test(raw_data$`Synchrony Index`) 
```

log

```{r}
raw_data.i <- subset(raw_data, raw_data$`Synchrony Index` != 0) #removing NA and values of 0
data_for_stats.8 <- raw_data.i %>% 
  drop_na(`Synchrony Index`) %>%
  mutate("LogParameter"=log(`Synchrony Index`))

model_treatment.8 <- lm(`LogParameter` ~ Treatment * Time , data = data_for_stats.8) 
 #diagnostic plots

Diagnostic_plot.8 <- autoplot(model_treatment.8, smooth.colour = NA) 
print(Diagnostic_plot.8)


anova(model_treatment.8)

summary(aov(model_treatment.8))
```

NO significance

fold change

```{r}
si_data <- raw_data %>%
  drop_na(`fold_change_si`)
```

```{r}
ggplot(si_data, aes(x = `Treatment`)) +
  geom_boxplot(aes(y = `fold_change_si`, colour = Treatment)) 
```

```{r}
ggplot(si_data, aes(x = `Treatment`, y = `fold_change_si`, colour = `Treatment`, fill = `Treatment`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `fold_change_si`, colour = Treatment), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `fold_change_si`, fill = `Treatment`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('fold change Synchrony Index')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Synchrony Index 72h treated')
```

```{r}
model_treatment.9 <- lm(`fold_change_si`~ Treatment , data = si_data) 
 #diagnostic plots

Diagnostic_plot.9 <- autoplot(model_treatment.9, smooth.colour = NA) 
print(Diagnostic_plot.9)


anova(model_treatment.9)

summary(aov(model_treatment.9))
```

no sig

# active electrodes

```{r}

raw_data <- subset(raw_data, raw_data$`Active Electrodes` != 0)

ggplot(raw_data, aes(x = `Treatment`)) +
  geom_point(aes(y = `Active Electrodes`, colour = Plate), position = position_jitter(width = .05)) 
```

N3 - has 8 wells active, N2 - has all wells active , N1 has 17 wells active - added jitter to see individual plots
