---
title: "29-32 div"
author: "Kara O'Driscoll"
date: "2024-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## opening files

```{r}
rm(list=objects())
library(tidyverse)
library(stringr)  
library(dplyr) 
library(ggfortify)
library(agricolae)
library(cowplot)

#looking at treatment from day of to 1 week ish after 
Path_to_working_directory <- "/Users/karaodriscoll/Documents/Lab Masters/div 29-32"

setwd(Path_to_working_directory) #sets directory to location of dataset
#opening dataset 
Name_of_dataset <- gsub(".*/","", Path_to_working_directory)

list_of_files <- list.files(pattern="*.csv")

raw_data <- read_csv(list_of_files, id = "file_name", show_col_types = FALSE)

raw_data$day <- as.numeric(str_extract(raw_data$file_name, '(?<=DIV)\\d+')) #obtain data by day

```

raincloud

```{r}
packages <- c("cowplot", "readr", "ggplot2", "dplyr", "lavaan", "Hmisc")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))
}

# Load packages ----
library(ggplot2)

# Defining the geom_flat_violin function ----
# Note: the below code modifies the
# existing github page by removing a parenthesis in line 50

"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}

geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
          setup_data = function(data, params) {
            data$width <- data$width %||%
              params$width %||% (resolution(data$x, FALSE) * 0.9)
            
            # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
            data %>%
              group_by(group) %>%
              mutate(
                ymin = min(y),
                ymax = max(y),
                xmin = x,
                xmax = x + width / 2
              )
          },
          
          draw_group = function(data, panel_scales, coord) {
            # Find the points for the line to go all the way around
            data <- transform(data,
                              xminv = x,
                              xmaxv = x + violinwidth * (xmax - x)
            )
            
            # Make sure it's sorted properly to draw the outline
            newdata <- rbind(
              plyr::arrange(transform(data, x = xminv), y),
              plyr::arrange(transform(data, x = xmaxv), -y)
            )
            
            # Close the polygon: set first and last point the same
            # Needed for coord_polar and such
            newdata <- rbind(newdata, newdata[1, ])
            
            ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
          },
          
          draw_key = draw_key_polygon,
          
          default_aes = aes(
            weight = 1, colour = "grey20", fill = "white", size = 0.5,
            alpha = NA, linetype = "solid"
          ),
          
          required_aes = c("x", "y")
  )
```

# number of spikes

```{r}
spikes_data <- raw_data %>%
  filter(Well.Averages=="Number of Spikes") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Number of Spikes")

#if cleared, need to rerun treatment df too
#treatment data frame 
treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")
#to join

spikes_with_treatment_data <- left_join(spikes_data, treatment, by = c("file_name", "day", "well_ID"))


```

```{r}
baseline_for_spikes <- spikes_with_treatment_data %>%
  filter(day==22) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(Baseline_spikes=`Number of Spikes`)

day29_for_spikes <- spikes_with_treatment_data %>%
  filter(day==29) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint1_spikes=`Number of Spikes`)

day30_for_spikes <- spikes_with_treatment_data %>%
  filter(day==30) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint2_spikes=`Number of Spikes`)

day31_for_spikes <- spikes_with_treatment_data %>%
  filter(day==31) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint3_spikes=`Number of Spikes`)

day32_for_spikes <- spikes_with_treatment_data %>%
  filter(day==32) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint4_spikes=`Number of Spikes`)

merged_data_spikes <- merge(baseline_for_spikes,day29_for_spikes, by = c("well_ID", "Treatment" ))
merged_data_spikes <- merge(merged_data_spikes, day30_for_spikes, by = c("well_ID", "Treatment"))
merged_data_spikes <- merge(merged_data_spikes, day31_for_spikes, by = c("well_ID", "Treatment"))
merged_data_spikes <- merge(merged_data_spikes, day32_for_spikes, by = c("well_ID", "Treatment"))

merged_data_normalised <- mutate(merged_data_spikes, "22"= Baseline_spikes - Baseline_spikes) 
merged_data_normalised. <- mutate(merged_data_normalised, "29"= timepoint1_spikes - Baseline_spikes) 
merged_data_normalised.. <- mutate(merged_data_normalised., "30"= timepoint2_spikes - Baseline_spikes)
merged_data_normalised... <- mutate(merged_data_normalised.., "31"= timepoint3_spikes - Baseline_spikes)
nmerged_data_normalised <- mutate(merged_data_normalised..., "32"= timepoint4_spikes - Baseline_spikes)

spikes_normalised <- nmerged_data_normalised %>%
  select(-c(3:7)) %>%
  pivot_longer(cols = c(3:7), 
               names_to = "Day", 
               values_to = "Normalised_data")
spikes_normalised$Day <- as.numeric(spikes_normalised$Day)
```

```{r}
spikes_normalised_avg <- spikes_normalised %>%
  group_by(Treatment, Day) %>% 
  mutate(average = mean(Normalised_data))
```

```{r}
spikes_normalised_avg <- spikes_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

ggplot(spikes_normalised_avg, aes(x = Day, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Number of Spikes by Treatment DIV 29-32",
       x = "Days in vitro (DIV)",
       y = "Normalised from baseline of Number of Spikes")+
  theme_cowplot()

ggsave("spikes_normalised_avg.png")
```

psil and 5meo inc but think doi all died :(

```{r}
spikes_normalised_avg$Treatment <- as.factor(spikes_normalised_avg$Treatment)
spikes_normalised_avg$Day <- as.factor(spikes_normalised_avg$Day)

data_for_stats <- spikes_normalised_avg %>% 
  drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~ Treatment * Day, data = data_for_stats) 
 #diagnostic plots

Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

# mean firing rate

```{r}
mean_firing_rate_data <- raw_data %>%
  filter(Well.Averages=="Mean Firing Rate (Hz)") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Mean Firing Rate (Hz)")

#if cleared, need to rerun treatment df too
#treatment data frame 
treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")
#to join

mean_firing_rate_with_treatment_data <- left_join(mean_firing_rate_data, treatment, by = c("file_name", "day", "well_ID"))

```

```{r}
baseline_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(day==22) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(Baseline_mfr=`Mean Firing Rate (Hz)`)

day29_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(day==29) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint1_mfr=`Mean Firing Rate (Hz)`)

day30_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(day==30) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint2_mfr=`Mean Firing Rate (Hz)`)

day31_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(day==31) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint3_mfr=`Mean Firing Rate (Hz)`)

day32_for_mfr <- mean_firing_rate_with_treatment_data %>%
  filter(day==32) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint4_mfr=`Mean Firing Rate (Hz)`)

merged_data_mfr <- merge(baseline_for_mfr,day29_for_mfr, by = c("well_ID", "Treatment"))
merged_data_mfr <- merge(merged_data_mfr, day30_for_mfr, by = c("well_ID", "Treatment"))
merged_data_mfr <- merge(merged_data_mfr, day31_for_mfr, by = c("well_ID", "Treatment"))
merged_data_mfr <- merge(merged_data_mfr, day32_for_mfr, by = c("well_ID", "Treatment"))

merged_data_normalised <- mutate(merged_data_mfr, "22"= Baseline_mfr - Baseline_mfr) 
merged_data_normalised. <- mutate(merged_data_normalised, "29"= timepoint1_mfr - Baseline_mfr) 
merged_data_normalised.. <- mutate(merged_data_normalised., "30"= timepoint2_mfr - Baseline_mfr)
merged_data_normalised... <- mutate(merged_data_normalised.., "31"= timepoint3_mfr - Baseline_mfr)
merged_data_normalised.... <- mutate(merged_data_normalised..., "32"= timepoint4_mfr - Baseline_mfr)


mfr_normalised <- merged_data_normalised.... %>%
  select(-c(3:7)) %>%
  pivot_longer(cols = c(3:7), 
               names_to = "Day", 
               values_to = "Normalised_data")
mfr_normalised$Day <- as.numeric(mfr_normalised$Day)
```

```{r}
mfr_normalised_avg <- mfr_normalised %>%
  group_by(Treatment, Day) %>% 
  mutate(average = mean(Normalised_data))
```

```{r}
mfr_normalised_avg <- mfr_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

ggplot(mfr_normalised_avg, aes(x = Day, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Mean Firing Rate by Treatment DIV 29-32",
       x = "Days in vitro (DIV)",
       y = "Normalised from baseline of Mean Firing Rate (Hz)")+
  theme_cowplot()

ggsave("mfr_normalised_avg.png")
```

```{r}
mfr_normalised_avg$Treatment <- as.factor(mfr_normalised_avg$Treatment)
mfr_normalised_avg$Day <- as.factor(mfr_normalised_avg$Day)

data_for_stats <- mfr_normalised_avg %>% 
  drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Day, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))

```

# burst freq

```{r}
burst_frequency <- raw_data %>%
  filter(Well.Averages=="Burst Frequency - Avg (Hz)") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26),
               names_to = "well_ID",
               values_to = "bursts")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#join
burst_freq_with_treatment <- left_join(burst_frequency, treatment, by = c("file_name", "day", "well_ID"))
```

```{r}
baseline_for_bf <- burst_freq_with_treatment %>%
  filter(day==22) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(Baseline_burst=`bursts`)

day29_for_bf <- burst_freq_with_treatment %>%
  filter(day==29) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint1_burst=`bursts`)

day30_for_bf <- burst_freq_with_treatment %>%
  filter(day==30) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint2_burst=`bursts`)

day31_for_bf <- burst_freq_with_treatment %>%
  filter(day==31) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint3_burst=`bursts`)

day32_for_bf <- burst_freq_with_treatment %>%
  filter(day==32) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint4_burst=`bursts`)

merged_data_bf <- merge(baseline_for_bf, day29_for_bf, by = c("well_ID", "Treatment"))
merged_data_bf <- merge(merged_data_bf, day30_for_bf, by = c("well_ID", "Treatment"))
merged_data_bf <- merge(merged_data_bf, day31_for_bf, by = c("well_ID", "Treatment"))
merged_data_bf <- merge(merged_data_bf, day32_for_bf, by = c("well_ID", "Treatment"))

bfmerged_data_normalised <- mutate(merged_data_bf, "22"= Baseline_burst - Baseline_burst) 
bfmerged_data_normalised. <- mutate(bfmerged_data_normalised, "29"= timepoint1_burst - Baseline_burst) 
bfmerged_data_normalised.. <- mutate(bfmerged_data_normalised., "30"= timepoint2_burst - Baseline_burst)
bfmerged_data_normalised... <- mutate(bfmerged_data_normalised.., "31"= timepoint3_burst - Baseline_burst)
bfmerged_data_normalised.... <- mutate(bfmerged_data_normalised..., "32"= timepoint4_burst - Baseline_burst)

bf_normalised <- bfmerged_data_normalised.... %>%
  select(-c(3:7)) %>%
  pivot_longer(cols = c(3:7), 
               names_to = "Day", 
               values_to = "Normalised_data")
bf_normalised$Day <- as.numeric(bf_normalised$Day)

#remove no data - after normalised otherewise it removed the parameters
bf_normalised <- na.omit(bf_normalised)
```

```{r}
bf_normalised_avg <- bf_normalised %>%
  group_by(Treatment, Day) %>% 
  mutate(average = mean(Normalised_data)
         )

```

```{r}
bf_normalised_avg <- bf_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))


ggplot(bf_normalised_avg, aes(x = Day, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Burst Frequency by Treatment DIV 29-32",
       x = "Days in vitro (DIV)",
       y = "Normalised from baseline of Burst Frequency (Hz)")+
  theme_cowplot()

ggsave("bf_normalised_avg.png")
```

not many wells

```{r}
bf_normalised_avg$Treatment <- as.factor(bf_normalised_avg$Treatment)
bf_normalised_avg$Day <- as.factor(bf_normalised_avg$Day)

data_for_stats <- bf_normalised_avg %>% 
 drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Day, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

no sig!

# spb

```{r}
spikes_per_burst <- raw_data %>%
  filter(Well.Averages=="Number of Spikes per Burst - Avg") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26),
               names_to = "well_ID",
               values_to = "bursts")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#join
spikes_per_burst_with_treatment <- left_join(spikes_per_burst, treatment, by = c("file_name", "day", "well_ID"))

```

```{r}
baseline_for_burst <- spikes_per_burst_with_treatment %>%
  filter(day==22) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(Baseline_burst=`bursts`)

day29_for_burst <- spikes_per_burst_with_treatment %>%
  filter(day==29) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint1_burst=`bursts`)

day30_for_burst <- spikes_per_burst_with_treatment %>%
  filter(day==30) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint2_burst=`bursts`)

day31_for_burst <- spikes_per_burst_with_treatment %>%
  filter(day==31) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint3_burst=`bursts`)

day32_for_burst <- spikes_per_burst_with_treatment %>%
  filter(day==32) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint4_burst=`bursts`)

merged_data_burst <- merge(baseline_for_burst, day29_for_burst, by = c("well_ID", "Treatment"))
merged_data_burst <- merge(merged_data_burst, day30_for_burst, by = c("well_ID", "Treatment" ))
merged_data_burst <- merge(merged_data_burst, day31_for_burst, by = c("well_ID", "Treatment" ))
merged_data_burst <- merge(merged_data_burst, day32_for_burst, by = c("well_ID", "Treatment" ))

bmerged_data_normalised <- mutate(merged_data_burst, "22"= Baseline_burst - Baseline_burst) 
bmerged_data_normalised. <- mutate(bmerged_data_normalised, "29"= timepoint1_burst - Baseline_burst) 
bmerged_data_normalised.. <- mutate(bmerged_data_normalised., "30"= timepoint2_burst - Baseline_burst)
bmerged_data_normalised... <- mutate(bmerged_data_normalised.., "31"= timepoint3_burst - Baseline_burst)
bmerged_data_normalised.... <- mutate(bmerged_data_normalised..., "32"= timepoint4_burst - Baseline_burst)

burst_normalised <- bmerged_data_normalised.... %>%
  select(-c(3:7)) %>%
  pivot_longer(cols = c(3:7), 
               names_to = "Day", 
               values_to = "Normalised_data")
burst_normalised$Day <- as.numeric(burst_normalised$Day)

#remove no data - after normalised otherewise it removed the parameters
burst_normalised <- na.omit(burst_normalised)
```

```{r}
burst_normalised_avg <- burst_normalised %>%
  group_by(Treatment, Day) %>% 
  mutate(average = mean(Normalised_data), IQR = IQR(Normalised_data))
```

```{r}
burst_normalised_avg <- burst_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))


ggplot(burst_normalised_avg, aes(x = Day, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Spikes per Burst by Treatment DIV 29-32",
       x = "Days in vitro (DIV)",
       y = "Normalised from baseline of Spikes per Burst")+
  theme_cowplot()

ggsave("burst_normalised_avg.png")
```

ummmmmm

```{r}
burst_normalised_avg$Treatment <- as.factor(burst_normalised_avg$Treatment)
burst_normalised_avg$Day <- as.factor(burst_normalised_avg$Day)

data_for_stats <- burst_normalised_avg %>% 
  drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Day, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

# synchrony

```{r}
synchrony <- raw_data %>%
  filter(Well.Averages=="Synchrony Index") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Synchrony Index")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#join
synchrony_with_treatment <- left_join(synchrony, treatment, by = c("file_name", "day", "well_ID"))
```

```{r}
baseline_for_si <- synchrony_with_treatment %>%
  filter(day==22) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(Baseline_si=`Synchrony Index`)

day29_for_si <- synchrony_with_treatment %>%
  filter(day==29) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint1_si=`Synchrony Index`)

day30_for_si <- synchrony_with_treatment %>%
  filter(day==30) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint2_si=`Synchrony Index`)

day31_for_si <- synchrony_with_treatment %>%
  filter(day==31) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint3_si=`Synchrony Index`)

day32_for_si <- synchrony_with_treatment %>%
  filter(day==32) %>%
  select(-c("file_name", "Well.Averages.x", "Well.Averages.y", "day")) %>%
  rename(timepoint4_si=`Synchrony Index`)

merged_data_si <- merge(baseline_for_si,day29_for_si, by = c("well_ID", "Treatment"))
merged_data_si <- merge(merged_data_si, day30_for_si, by = c("well_ID", "Treatment"))
merged_data_si <- merge(merged_data_si, day31_for_si, by = c("well_ID", "Treatment"))
merged_data_si <- merge(merged_data_si, day32_for_si, by = c("well_ID", "Treatment"))

smerged_data_normalised <- mutate(merged_data_si, "22"= Baseline_si - Baseline_si) 
smerged_data_normalised. <- mutate(smerged_data_normalised, "29"= timepoint1_si - Baseline_si) 
smerged_data_normalised.. <- mutate(smerged_data_normalised., "30"= timepoint2_si - Baseline_si)
smerged_data_normalised... <- mutate(smerged_data_normalised.., "31"= timepoint3_si - Baseline_si)
smerged_data_normalised.... <- mutate(smerged_data_normalised..., "32"= timepoint4_si - Baseline_si)

si_normalised <- smerged_data_normalised.... %>%
  select(-c(3:7)) %>%
  pivot_longer(cols = c(3:7), 
               names_to = "Day", 
               values_to = "Normalised_data")
si_normalised$Day <- as.numeric(si_normalised$Day)

si_normalised <- na.omit(si_normalised)
```

```{r}
si_normalised_avg <- si_normalised %>%
  group_by(Treatment, Day) %>% 
  mutate(average = mean(Normalised_data), IQR = IQR(Normalised_data))
```

```{r}
si_normalised_avg <- si_normalised_avg %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))



ggplot(si_normalised_avg, aes(x = Day, colour = Treatment)) +
  geom_line(aes(y=average)) +
  geom_point(aes(y=Normalised_data, shape = Treatment)) +
  labs(title = "Normalised Sychrony Index by Treatment DIV 29-32",
       x = "Days in vitro (DIV)",
       y = "Normalised from baseline of Synchrony Index")+
  theme_cowplot()
```

```{r}
si_normalised_avg$Treatment <- as.factor(si_normalised_avg$Treatment)
si_normalised_avg$Day <- as.factor(si_normalised_avg$Day)

data_for_stats <- si_normalised_avg %>% 
 drop_na(`Normalised_data`) %>%
  mutate("LogParameter"=log(`Normalised_data`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*Day, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

```{r}
TukeyHSD(aov(model_treatment), "Treatment", ordered = TRUE)

summary(aov(model_treatment), test = adjusted(type = "BH"))
```

yeah only sig is to DOI CHANGING RAHHHHH

anywya

# netwerk spb

```{r}

net_spikes_per_burst <- raw_data %>%
  filter(Well.Averages=="Number of Spikes per Network Burst - Avg") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26),
               names_to = "well_ID",
               values_to = "bursts")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#join
net_spikes_per_burst_with_treatment <- left_join(net_spikes_per_burst, treatment, by = c("file_name", "day", "well_ID"))

```

```{r}
net_spikes_per_burst_with_treatment <- net_spikes_per_burst_with_treatment %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

net_spikes_per_burst_with_treatment$day <- as.factor(net_spikes_per_burst_with_treatment$day)


ggplot(net_spikes_per_burst_with_treatment, aes(x = `Treatment`, y = `bursts`, colour = `day`, fill = `day`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `bursts`, colour = `day`), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `bursts`, fill = `day`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('Spikes per Network Burst')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Spikes per Network Burst DIV 29-32')

ggsave("net_spikes_per_burst_with_treatment.png")
```

kind of shows but quite badlyyyyyy - the inc in network bursts as days go on!!!

```{r}
shapiro.test(net_spikes_per_burst_with_treatment$bursts) #is normally dis

net_spikes_per_burst_with_treatment$Treatment <- as.factor(net_spikes_per_burst_with_treatment$Treatment)
net_spikes_per_burst_with_treatment$day <- as.factor(net_spikes_per_burst_with_treatment$day)


model_treatment <- lm(`bursts`~Treatment*day, data = net_spikes_per_burst_with_treatment) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)


anova(model_treatment)

summary(aov(model_treatment))
```

treatment sig UGH GOING TO BE DOI

```{r}
TukeyHSD(aov(model_treatment), "Treatment", ordered = TRUE)

summary(aov(model_treatment), test = adjusted(type = "BH"))
```

OHHHH NO THERE IS SIG COMPARED TO VEHICLE WOOOOOO

# netwerk burst frequency

```{r}
network_burst_frequency <- raw_data %>%
  filter(Well.Averages=="Network Burst Frequency (Hz)") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26),
               names_to = "well_ID",
               values_to = "burstz")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

#join
net_burst_freq_with_treatment <- left_join(network_burst_frequency, treatment, by = c("file_name", "day", "well_ID"))
```

```{r}
net_burst_freq_with_treatment <- net_burst_freq_with_treatment %>%
  mutate(Treatment = fct_relevel(Treatment, "Vehicle", "5-MEO-DMT [0.5 µM]", "Psilocin [1 µM]", "DOI [10 µM]"))

net_burst_freq_with_treatment$day <- as.factor(net_burst_freq_with_treatment$day)


ggplot(net_burst_freq_with_treatment, aes(x = `Treatment`, y = `burstz`, colour = `day`, fill = `day`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `burstz`, colour = `day`), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `burstz`, fill = `day`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('Network Burst Frequency (Hz)')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Network Burst Frequency (Hz) DIV 29-32')

ggsave("net_burst_freq_with_treatment.png")
```

ahhhhhh not really as exciting

```{r}
shapiro.test(net_burst_freq_with_treatment$burstz) #not normal = looooogggg

net_burst_freq_with_treatment$Treatment <- as.factor(net_burst_freq_with_treatment$Treatment)
net_burst_freq_with_treatment$day <- as.factor(net_burst_freq_with_treatment$day)

data_for_stats <- net_burst_freq_with_treatment %>% 
 drop_na(`burstz`) %>%
  mutate("LogParameter"=log(`burstz`)) %>%
filter(is.finite(LogParameter))

model_treatment <- lm(`LogParameter`~Treatment*day, data = data_for_stats) 
 #diagnostic plots
Diagnostic_plot <- autoplot(model_treatment, smooth.colour = NA) 
print(Diagnostic_plot)

anova(model_treatment)

summary(aov(model_treatment))
```

no sig

```{r}
ggplot(data_for_stats, aes(x = `Treatment`, y = `LogParameter`, colour = `day`, fill = `day`))+
   geom_flat_violin(position = position_nudge(x = .1, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) + #adjust function adjusts the smoothness - make snmaller for less of a smooth curve 
  geom_point(aes(y = `LogParameter`, colour = `day`), position = position_jitter(width = .15), size = 1, shape = 20)+ 
 geom_boxplot(aes(x = `Treatment`, y = `LogParameter`, fill = `day`), outlier.shape = NA, alpha = .5, width = .1, colour = "black") +
  ylab('log Network Burst Frequency (Hz)')+
  xlab('Treatment')+
  coord_flip()+
  theme_cowplot()+
  guides(fill = FALSE)+
  ggtitle('Network Burst Frequency (Hz) DIV 29-32')
```

looks also bad! SHAME
