---
title: "pre treatment all plates div22"
author: "Kara O'Driscoll"
date: "2024-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=objects())
library(tidyverse)
library(stringr)  
library(dplyr) 
library(ggfortify)
library(cowplot)
library(readr)
library(RColorBrewer)


Path_to_working_directory <- "/Users/karaodriscoll/Documents/Lab Masters/all plates just before treatment"
setwd(Path_to_working_directory) #sets directory to location of dataset
#opening dataset 
Name_of_dataset <- gsub(".*/","", Path_to_working_directory)
list_of_files <- list.files(pattern="*.csv")
raw_data <- read_csv(list_of_files, id = "file_name", show_col_types = FALSE)

raw_data$plate <- as.numeric(str_extract(raw_data$file_name, "(?<=N)\\d+"))
raw_data$plate <- as.factor(raw_data$plate)

```

```{r}
##adding in rainplot - from https://github.com/RainCloudPlots

packages <- c("cowplot", "readr", "ggplot2", "dplyr", "lavaan", "Hmisc")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))
}

# Load packages ----
library(ggplot2)

# Defining the geom_flat_violin function ----
# Note: the below code modifies the
# existing github page by removing a parenthesis in line 50

"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}

geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
          setup_data = function(data, params) {
            data$width <- data$width %||%
              params$width %||% (resolution(data$x, FALSE) * 0.9)
            
            # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
            data %>%
              group_by(group) %>%
              mutate(
                ymin = min(y),
                ymax = max(y),
                xmin = x,
                xmax = x + width / 2
              )
          },
          
          draw_group = function(data, panel_scales, coord) {
            # Find the points for the line to go all the way around
            data <- transform(data,
                              xminv = x,
                              xmaxv = x + violinwidth * (xmax - x)
            )
            
            # Make sure it's sorted properly to draw the outline
            newdata <- rbind(
              plyr::arrange(transform(data, x = xminv), y),
              plyr::arrange(transform(data, x = xmaxv), -y)
            )
            
            # Close the polygon: set first and last point the same
            # Needed for coord_polar and such
            newdata <- rbind(newdata, newdata[1, ])
            
            ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
          },
          
          draw_key = draw_key_polygon,
          
          default_aes = aes(
            weight = 1, colour = "grey20", fill = "white", size = 0.5,
            alpha = NA, linetype = "solid"
          ),
          
          required_aes = c("x", "y")
  )
```

```{r}
active_electrodes <- raw_data %>%
   filter(Well.Averages=="Number of Active Electrodes") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Active_Electrodes")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

electrodes_with_treatment_data <- left_join(active_electrodes, treatment, by = c("file_name", "plate", "well_ID"))

electrodes_with_treatment_data$Active_Electrodes <- as.numeric(electrodes_with_treatment_data$Active_Electrodes)

electrodes_with_treatment_data <- subset(electrodes_with_treatment_data, electrodes_with_treatment_data$Active_Electrodes != 0) #removing inavtive electrodes

```

plotting the number of active electrodes

```{r}
ggplot(electrodes_with_treatment_data, aes(x = `Treatment`)) +
  geom_point(aes(y = `Active_Electrodes`, colour = plate), position = position_jitter(width = 0.15)) +
  theme_cowplot()+
  ylab('Active Electrodes')+
  xlab('Treatment')+
  ggtitle('Active Electrodes before Treatment')
  
```

9 wells active - n3 - one well 5meo, 4 wells doi, 2 wells psilocin, 2 wells vehicle

19 wells active - n1 - 5 wells 5meo, 5 wells doi, 4 wells psilocin, 5 wells vehicle

24 wells active - n2 - 6 for all

# number of spikes

extracting the number of spikes and treatment

```{r}
number_of_spikes_data <- raw_data %>%
  filter(Well.Averages=="Number of Spikes") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Spikes")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

spikes_with_treatment_data <- left_join(number_of_spikes_data, treatment, by = c("file_name", "plate", "well_ID"))

spikes_with_treatment_data <- subset(spikes_with_treatment_data, spikes_with_treatment_data$Spikes != 0)

spikes_avg <- spikes_with_treatment_data %>%
    group_by(Treatment, plate) %>% 
  mutate(average = mean(Spikes))

```

plotting the raw data

```{r}
ggplot(spikes_with_treatment_data, aes(x = Treatment))+ 
  geom_point(aes( y = Spikes, colour = plate),position = position_jitter(width = 0.10) )


```

plotting the averaged data = to do this I averaged across wells per condition per plate

```{r}
ggplot(spikes_avg, aes(x = Treatment))+ 
  geom_point(aes( y = average, colour = plate))

```

# mean firing rate

extracting mean firing rate

```{r}
mean_firing_rate_data <- raw_data %>%
  filter(Well.Averages=="Mean Firing Rate (Hz)") %>%
  mutate(across(-c(1,2,27), as.numeric, na.rm = TRUE)) %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "mfr")

treatment <- raw_data %>%
  filter(Well.Averages=="Treatment/ID") %>%
  pivot_longer(cols = c(3:26), 
               names_to = "well_ID", 
               values_to = "Treatment")

mfr_with_treatment_data <- left_join(mean_firing_rate_data, treatment, by = c("file_name", "plate", "well_ID"))

mfr_with_treatment_data <- subset(mfr_with_treatment_data, mfr_with_treatment_data$mfr != 0)

mfr_avg <- mfr_with_treatment_data %>%
  group_by(Treatment, plate) %>% 
  mutate(average = mean(mfr))
```

plotting mean firing rate raw data

```{r}
ggplot(mfr_avg, aes(x = Treatment))+ 
  geom_point(aes( y = mfr, colour = plate)) 
```

plotting mean firing rate averaged data

```{r}
ggplot(mfr_avg, aes(x = Treatment)) +
  geom_point(aes(y = average, colour = plate))
```
